Random_Effects

### 2023-07-09 Code/BatchEffect.Rmd
Renamed several variables and fixed the commenting issues

```{r}
times <- read.csv("../Data/ThesisData.csv")
colnames(times) <- c("Year", "Type", "Time", "RxnTime", "Gender", "Batch")
men_finals <- subset(times, Type == "F" & Gender == "M" & Time != 'DQ')
men_pooled <- subset(times, (Type == "F" | Type == "S") & 
                       Gender == "M" & Time != 'DQ')
men_old_finals <- subset(times, Type == "F" & Gender == "M" & 
                           Time != 'DQ' & Year < 2022)
men_old_pooled <- subset(times, (Type == "F" | Type == "S") & 
                           Gender == "M" & Time != 'DQ' & Year < 2022)
```
```{}
This loads in the datasets
```

@@

```
```{}
This creates the year effect and the no year effect models
```

```{r}
men_old_finals_batch <-glmer(RxnTime ~ (1 | Batch), data = old_finals, 
                             family = Gamma(link = "log"))
men_finals_batch <-glmer(RxnTime ~ (1 | Batch), data = men_finals, 
                         family = Gamma(link = "log"))
men_old_pooled_batch <-glmer(RxnTime ~ (1 | Batch), data = old_pooled, 
                             family = Gamma(link = "log"))
men_pooled_batch <-glmer(RxnTime ~ (1 | Batch), data = men_pooled, 
                         family = Gamma(link = "log"))
```
```{}
This creates the batch effect only models
```

```{r}
men_old_finals_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
                                   data = old_finals, family = Gamma(link = "log"))
men_finals_two_effect<- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
                              data = men_finals, family = Gamma(link = "log"))
men_old_pooled_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
                                   data = old_pooled, family = Gamma(link = "log"))
men_pooled_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
                               data = men_pooled, family = Gamma(link = "log"))
```
```{}
This creates the year and batch effect models
```

```{r}
anova(men_pooled_batch, men_pooled_two_effect)
```
```{}
This anova test determined that the two_effect model is an improvement over the
batch effect model. The test statistic is significant at alpha = 0.05 but not at
alpha = 0.01.
```

```{r}
anova(men_pooled_year_gamma_log, men_pooled_two_effect)
```
```{}
In this anova test, it is clear that the two random effect model is superior to
having only the year effect.
```

@@

  sbase <- summary(base)
  x <- simfit(model)
  print(mean(x < RxnTime))
  #print(quantile(x, prob = c(0.00001, 0.0001, 0.0005, 0.001)))
  ## compare with and without random effect
  plot(density(x))  # with random effects
  curve(dgamma(x, shape = 1 / sbase$dispersion, 
               scale = exp(coef(base)) * sbase$dispersion), 
        0.08, 0.22, add = TRUE, col = "blue")  # fixed effect only
}
simulation_two_effect(old_finals_two_effect, old_finals_one_gamma_log, 0.1)
simulation_two_effect(men_finals_two_effect, men_finals_one_gamma_log, 0.1)
simulation_two_effect(old_pooled_two_effect, old_pooled_one_gamma_log, 0.1)
simulation_two_effect(men_pooled_two_effect, men_pooled_one_gamma_log, 0.1)
```

```{}
This code simulates the probability of observing a reaction time less than 0.1
for the various datasets.
```

>>>>

```{r}
times <- read.csv("../Data/ThesisData.csv")
colnames(times) <- c("Year", "Type", "Time", "RxnTime", "Gender", "Batch")
men_finals <- subset(times, Type == "F" & Gender == "M" & Time != 'DQ')
men_pooled <- subset(times, (Type == "F" | Type == "S") & 
                       Gender == "M" & Time != 'DQ')
men_old_finals <- subset(times, Type == "F" & Gender == "M" & 
                           Time != 'DQ' & Year < 2022)
men_old_pooled <- subset(times, (Type == "F" | Type == "S") & 
                           Gender == "M" & Time != 'DQ' & Year < 2022)
```
This loads in the 4 datasets from the csv file.

@@

```
This creates the year effect and the no effect models.  The models without a year
effect end in "one_gamma_log" as in men_old_finals_one_gamma_log.

```{r}
men_old_finals_batch_gamma_log <-glmer(RxnTime ~ (1 | Batch), data = old_finals, 
                             family = Gamma(link = "log"))
men_finals_batch_gamma_log <-glmer(RxnTime ~ (1 | Batch), data = men_finals, 
                         family = Gamma(link = "log"))
men_old_pooled_batch_gamma_log <-glmer(RxnTime ~ (1 | Batch), data = old_pooled, 
                             family = Gamma(link = "log"))
men_pooled_batch_gamma_log <-glmer(RxnTime ~ (1 | Batch), data = men_pooled, 
                         family = Gamma(link = "log"))
```

This creates the models that have the batch effect instead of the year effect.

```{r}
men_old_finals_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
                                   data = old_finals, family = Gamma(link = "log"))
men_finals_two_effect_gamma_log<- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
                              data = men_finals, family = Gamma(link = "log"))
men_old_pooled_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
                                   data = old_pooled, family = Gamma(link = "log"))
men_pooled_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
                               data = men_pooled, family = Gamma(link = "log"))
```
This creates the models with both a year and batch effect.

```{r}
anova(men_pooled_batch, men_pooled_two_effect)
```

This anova test determined that the two effect model is an improvement over the
batch effect model. The test statistic is 0.01059 which is significant at alpha 
= 0.05 but not at alpha = 0.01.

```{r}
anova(men_pooled_year_gamma_log, men_pooled_two_effect)
```
This anova test determined that the two effect model is a significant improvement
over the year effect model.

@@

  sbase <- summary(base)
  x <- simfit(model)
  print(mean(x < RxnTime))
  #print(quantile(x, prob = c(0.00001, 0.0001, 0.0005, 0.001)))
  ## compare with and without random effect
  plot(density(x))  # with random effects
  curve(dgamma(x, shape = 1 / sbase$dispersion, 
               scale = exp(coef(base)) * sbase$dispersion), 
        0.08, 0.22, add = TRUE, col = "blue")  # fixed effect only
}
simulation_two_effect(old_finals_two_effect, old_finals_one_gamma_log, 0.1)
simulation_two_effect(men_finals_two_effect, men_finals_one_gamma_log, 0.1)
simulation_two_effect(old_pooled_two_effect, old_pooled_one_gamma_log, 0.1)
simulation_two_effect(men_pooled_two_effect, men_pooled_one_gamma_log, 0.1)
```
This code simulates the probability of observing a reaction time less than 0.1
for the various data sets.  The function takes in two models and a reaction time,
it is important that the two models that are compared are from the same data set;
we want to compare old_finals_two_effect and old_finals_one_gamma_log.
